Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative genomics

### 2018-02-28 paper-webserver.tex
Added captions to the figures

\mwcomment{I would suggest we avoid the comparison to sequence databases or genome sequence platforms (e.g. assembly, gene annotation etc). We really don't have the capacity to replace EnteroBase, Genbank, IRIDA, etc. I see Spfy as a repository of E.coli phenotypes important for public health. It allows users to rapidly predict phenotypes and retrieve the phenotypes of historical strains to: track trends, observe context amoung evolutionary related strains, identify biomarkers, etc. In other words, Spfy provides WGS-based phenotypes + big picture comparative genomics. The big picture is only possible by 1. recomputing everything everytime, or 2. storing the results as genomes are analysed. With size of WGS data, option 2 is only viable option}
\mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
\title{Spfy: an integrated graph database and analysis platform for real-time prediction of Escherichia coli phenotypes}

@@

Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue. Results are converted into individual graphs and stored within a larger graph database according to standard ontologies {list}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.

@@

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}. For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}. This system is flexible and allows novel data to be incorporated into the existing graph.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{An example of how data is stored in Spfy. Brackets highlight the source of different data points and the software it was generated from. These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database.}
\label{fig-ontology}
\end{figure}

@@

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/results.png}
\end{center}
\caption{Caption for figure within column.}
\label{fig-results}
\end{figure}

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/tables.png}
\end{center}
\caption{Caption for figure within column.}
\label{fig-tables}
\end{figure}

@@

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/docker}
\end{center}
\caption{Caption for figure within column.}
\label{fig-docker}
\end{figure}

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational inferface which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed. The Galaxy worklow focus is on running an individual analysis pipeline. In does not include functionality to store and collate analysis results for large-scale comparative analysis.

>>>>

\mwcomment{I would suggest we avoid the comparison to sequence databases or genome sequence platforms (e.g. assembly, gene annotation etc). We really don't have the capacity to replace EnteroBase, Genbank, IRIDA, etc. I see Spfy as a repository of E.coli phenotypes important for public health. It allows users to rapidly predict phenotypes and retrieve the phenotypes of historical strains to: track trends, observe context amoung evolutionary related strains, identify biomarkers, etc. In other words, Spfy provides WGS-based phenotypes + big picture comparative genomics. The big picture is only possible by 1. recomputing everything everytime, or 2. storing the results as genomes are analysed. With size of WGS data, option 2 is only viable option}
\mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
\title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative genomics}

@@

Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue. Results are converted into individual graphs and stored within a larger graph database according to the standard ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.

@@

\subsection{Data Storage}
% para
% 0. Goals: big-data, everything linked, easy addition of new links
% 1. spfy is built around graph technologies
% 1. how we structure our graph
% 2. ontoogies used
% 3. inferencing
% 3 1/2. SPARQL queries
% para: the semantic web
Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}. For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from given ontologies. This system is flexible and allows novel data to be incorporated into the existing graph.
\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/ontology}
\end{center}
\caption{Structure of Spfy's graph database. Brackets highlight the source of different data points and the software it was generated from. These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database. Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
\label{fig-ontology}
\end{figure}

@@

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/results.png}
\end{center}
\caption{First entry into the results interface for submitted tasks. Cards represent individual tasks and the entire collection can be viewed by sharing the URL with the embedded token.}
\label{fig-results}
\end{figure}

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/tables.png}
\end{center}
\caption{An expanded result for a subtyping task. While data storage in Spfy is graph-based, a familiar tabular structure is presented to users. The result can also be shared using the URL with the embedded token.}
\label{fig-tables}
\end{figure}

@@

\begin{figure}[!hb]
\begin{center}
\includegraphics[width=\textwidth]{images/docker}
\end{center}
\caption{The various Docker containers used in Spfy. Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using Docker-Compose.}
\label{fig-docker}
\end{figure}

@@

% namely galaxy
The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}. Galaxy aims to provide a reproducible, computational inferface which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed. The Galaxy worklow focus is on running an individual analysis pipeline. It does not include functionality to store and collate analysis results for large-scale comparative analysis.
